### 2020-03-11 paper/supporting_information.tex
working on suplement

\documentclass[10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[font=small,labelfont=bf]{caption}
\usepackage{geometry}
\usepackage{natbib}
\usepackage{hyperref}
\usepackage{pxfonts}
\usepackage{graphicx}
\usepackage{newfloat}
\usepackage{setspace}
%\doublespacing

\newcommand{\argmax}{\mathop{\mathrm{argmax}}\limits}

\title{\textit{Supplemental figures for}: A Gaussian process model of human electrocorticographic data}
\author{
  Lucy L. W. Owen$^{1}$,
  Tudor A. Muntianu$^{1}$,
  Andrew C. Heusser$^{1, 2}$, \\
  Patrick Daly$^{3}$,
  Katherine Scangos$^{3}$, and
  Jeremy R. Manning$^{1\ast}$\\\\
$^{1}$Department of Psychological and Brain Sciences, Dartmouth College,\\
Hanover, NH 03755, USA\\
$^{2}$Akili Interactive,\\
Boston, MA 02110, USA}

%\bibliographystyle{apa}

\begin{document}

\setcounter{equation}{0}
\setcounter{figure}{0}
\setcounter{table}{0}
\setcounter{page}{1}
\setcounter{section}{0}
\makeatletter
\renewcommand{\theequation}{S\arabic{equation}}
\renewcommand{\thefigure}{S\arabic{figure}}

\newcommand{\methods}{1}
\newcommand{\corrmaps}{2}
\newcommand{\freqs}{3}
\newcommand{\density}{4}
\newcommand{\infomap}{5}
\newcommand{\infomapfreqs}{6}

\begin{titlepage}
  \maketitle
\end{titlepage}

\begin{figure}[p]
\centering
\includegraphics[width=\textwidth]{figs/supplemental_1}
\caption{\textbf{Reconstruction accuracy across all electrodes in two
    ECoG datasets, broken down by experiment.}
  \textbf{A. Distributions of correlations between observed versus
    reconstructed activity by electrode, for Dataset 1.}  The
  across-patient distribution (black) reflects reconstruction accuracy
  (correlation) using a correlation model learned from all but one
  patient's data, and then applied to that held-out patient's data.
  The within-patient distribution (gray) reflects performance using a
  correlation model learned from the same patient who contributed the
  to-be-reconstructed electrode.  The split violin plot displays the
  same distributions shown in Figure~\corrmaps A.
  \textbf{B. Distributions of correlations for Dataset 2.}  The split
  violin plots are in the same format as those in Panel A.  The
  leftmost plot (``Within'') displays the same distributions shown in
  Figure~\corrmaps B.  All reconstructions reflected in the
  distribution were carried out using a model trained and tested using
  data from the same experiment.  The middle plots (``Across'')
  reflect reconstructions trained and tested using data from different
  experiments.  The rightmost plot (``All'') reflect reconstructions
  obtained using models trained and tested on data from both
  experiments.  The black distributions reflect models trained and
  tested across patients (analogous to the black histograms in
  Figure~\corrmaps) and the gray distributions reflect models trained
  and tested within patient (analogous to the gray histograms in
  Figure~\corrmaps).  The dark gray significance bars reflect
  paired-sample $t$-tests comparing the ($z$-transformed)
  reconstruction accuracy for each electrode obtained within versus
  across patients.  The black significance bars reflect paired-sample
  $t$-tests comparing the across-patient reconstruction accuracies
  across datasets.  The light gray significance bars reflect
  paired-sample $t$-tests comparing the within-patient reconstruction
  accuracies across datasets.  The symbols denote the corresponding
  $p$-values of those statistical tests.}
\label{fig:supplemental_1}
\end{figure}

\begin{figure}[p]
\centering
\includegraphics[width=\textwidth]{figs/supplemental_2}
\caption{\textbf{Reconstruction accuracy for Dataset 2, Experiments 1
    and 2.}  \textbf{A. Distributions of correlations between observed
    versus reconstructed activity by electrode, for Dataset 2
    (Experiment 1).}  The plots are in the same format as
  Figure~\ref{fig:supplemental_1}B, but reflect data only from
  Experiment 1 in Dataset 2.  \textbf{A. Distributions of correlations between observed
    versus reconstructed activity by electrode, for Dataset 2
    (Experiment 2).}  The plots are in the same format as Panel A, but
    reflect data only from Experiment 2 in Dataset 2.}
\label{fig:supplemental_2}
\end{figure}

\begin{figure}[p]
\centering \includegraphics[width=\textwidth]{figs/supplemental_3}
\caption{\textbf{Reconstruction accuracy across all electrodes in Dataset 2.}
This figure is analogous to Figure~\corrmaps in the main text, but displays data
for Dataset 2 broken down by experiment. \textbf{A. Distributions of
correlations between observed versus reconstructed activity by electrode for
Experiment 1.}  The across-patient distribution (black) reflects reconstruction
accuracy (correlation) using a correlation model learned from all but one
patient's data, and then applied to that held-out patient's data.  The
within-patient distribution (gray) reflects performance using a correlation
model learned from the same patient who contributed the to-be-reconstructed
electrode. \textbf{B. Distributions of correlations for Experiment 2.}  This
panel is in the same format as Panel A, but reflects results obtained from
Experiment 2. \textbf{C.--D. Reconstruction accuracy by location.} The colors
denote the average across-session correlations, using the across-patient
correlation model, between the observed and reconstructed activity at the given
electrode location projected to the cortical surface~\citep{CombEtal19}.  Panel
C displays the map for Experiment 1 and Panel D displays the map for Experiment 2.}
\label{fig:supplemental_3}
\end{figure}

\begin{figure}[p]
\centering
\includegraphics[width=\textwidth]{figs/supplemental_4}
\caption{\textbf{Reconstruction accuracy across all electrodes in two ECoG
datasets for power in each frequency band.} This figure is analogous to
Figure~\freqs in the main text, but displays the reconstruction accuracies of
the \textit{power} in each frequency band, rather than the bandpass-filtered
voltages, as in Figure~\freqs. \textbf{A. Distributions of correlations between
observed versus reconstructed power by electrode for each frequency band in
Dataset 1.}  Each color denotes a different frequency band. Within each color
group, the darker dots and bar on the left display the distribution (and mean)
across-patient reconstruction accuracies (analogous to the black histograms in
Fig.~\corrmaps).  The lighter dots and bar on the right display the distribution
(and mean) within-patient reconstruction accuracies (analogous to the gray
histograms in Fig.~\corrmaps). Each dot indicates the reconstruction accuracy
for one electrode in the dataset. \textbf{B.~Statistical summary of
across-patient reconstruction accuracy by electrode for each frequency band in
Dataset 1.} In the upper triangles of each map, warmer colors (positive
$t$-values) indicate that the reconstruction accuracy for the frequency band in
the given row was greater (via a two-tailed paired-sample $t$-test) than for the
frequency band in the given column. Cooler colors (negative $t$-values) indicate
that reconstruction accuracy for the frequency band in the given row was lower
than for the frequency band in the given column. The lower triangles of each map
denote the corresponding $p$-values for the $t$-tests. The diagonal entries
display the average reconstruction accuracy within each frequency band.
\textbf{C.~Statistical summary of within-patient reconstruction accuracy by
electrode for each frequency band in Dataset 1.} This panel displays the
within-patient statistical summary, in the same format as Panel B.
\textbf{D.~Distributions of correlations between observed versus reconstructed
activity by electrode, for each frequency band in Dataset 2.} This panel
displays reconstruction accuracy distributions for each frequency band for
Dataset 2. \textbf{E.--F.~Statistical summaries of across-patient and
within-patient reconstruction accuracy by electrode for each frequency band in
Dataset 2.} These panels are in the same as Panels B and C, but display results
from Dataset 2.}
\label{fig:supplemental_4}
\end{figure}

\begin{figure}[p]
\centering \includegraphics[width=\textwidth]{figs/supplemental_5}
\caption{\textbf{Most informative recording locations by frequency band.} This
figure is analogous to Figure~\infomapfreqs in the main text, but displays the
brain regions that were most informative about the \textit{power} in each
frequency band (whereas Fig.~\infomapfreqs displays the regions that were most
informative about the bandpass-filtered voltages within each band).
\textbf{A.~Intersections between information score maps by frequency band.} The
regions indicated in each row depict the intersection between the top 10\% most
informative locations across Datasets 1 and 2.  \textbf{B.~Network memberships
of the most informative brain regions.}  The pie charts display the proportions
of voxels in each region that belong to the seven networks identified by
\cite{YeoEtal11}.  The relative sizes of the charts for each frequency band
reflect the average across-subject reconstruction accuracies
(Figs.~\ref{fig:frequency}A, D).  The voxels in Panel A are colored according to
the same network memberships.  \textbf{C.~Neurosynth terms associated with the
most informative brain regions, by frequency band.}  The lists in each row
display the top five neurosynth terms~\citep{RubiEtal17} decoded for each
region.  \textbf{D.~Network parcellation map and legend.}  The parcellation
defined by \cite{YeoEtal11} is displayed on the inflated brain maps.  The colors
and network labels serve as a legend for Panels A and B. \textbf{E.~Combined map
of the most informative brain regions.}  The map displays the union of the most
informative maps in Panel A, colored by frequency band.  The labels also serve
as a legend for Panel C.}
\label{fig:supplemental_5}
\end{figure}

\begin{figure}[p]
\centering \includegraphics[width=\textwidth]{figs/supplemental_6}
\caption{\textbf{Reconstruction accuracy versus within-subject data features for
two ECoG datasets.} The individual dots in each panel  reflect data from one
patient.  The least-squares linear regression lines (with shaded 95\% confidence
intervals) and the correlations reported in each panel are between the data
shown on the $x$- and $y$-axes of the panel.  The $y$-axis in each panel
denotes the average $z$-transformed correlation between the observed and
predicted voltage timeseries at each electrode (across all of the patient's
electrodes). The cross-validated predictions were obtained using the
across-patient model (Fig.~\corrmaps A and B, black histograms).  \textbf{A.--F.
Dataset 1.}  \textbf{A. Total electrodes.}  The $x$-coordinates of each dot
display the total number of electrodes implanted in each patient's brain.
\textbf{B. Number thresholded.}  The $x$-coordinates of each dot display the
number of implanted electrodes that survived the kurtosis-based filtering
procedure (see \textit{Approach}).  \textbf{C. Proportion thresholded.} The
$x$-coordinates of each dot display the \textit{proportion} of each patient's
implanted electrodes (relative to the total number) that survived the
kurtosis-based filtering procedure.  \textbf{D. Sessions.}  The $x$-coordinates of each dot display the
number of distinct recording sessions contributed by each patient.  \textbf{E. Total time.} The $x$-coordinates of each dot display the
total duration (in hours) of each patient's recordings, across all of their recording sessions.  \textbf{F. Average session length.} The $x$-coordinates of each dot display average duration (in hours) of the patient's recording sessions.
  \textbf{G.--L. Features from Dataset 2.}  These panels display analogous information to Panels A--F, but for Dataset 2 patients.}
\label{fig:supplemental_6}
\end{figure}

%JRM STOPPED HERE
\begin{figure}[p]
\centering
\includegraphics[width=\textwidth]{figs/supplemental_7}
\caption{\textbf{Stability of correlation matrix across time.} \textbf{A.--B.  Features from
    Dataset 1.}  Features include: (\textbf{A. Correlation of split
  halves by the percent samples included.})  For each patient, we
  split the timeseries data in
  half and created a correlation matrix with one half.  With the
  remaining data, we created the second correlation matrix with more
  and more time samples (in 1000 sample increments).  These time
  samples were either drawn randomly (``Random'') , or were drawn sequentially from
  either the closest timepoint (``Close'' ) or the furthest timepoint
  (``Apart'' ). The two correlation matrices were then correlated, and the
  coefficient plotted relative to the percentage of data included.  (\textbf{B. Correlation of split
  halves by the number of time samples included.})
Same analysis as Panel A, but the
  coefficient plotted relative to the total number of samples included.
  \textbf{C.--D. Features from Dataset 2.}  Analogous format to Panels
  A--B.}
\label{fig:supplemental_7}
\end{figure}

\begin{figure}[p]
\centering
\includegraphics[width=\textwidth]{figs/supplemental_8}
\caption{\textbf{Stability of correlation matrix across patients.} \textbf{A.--B.  Features from
    Dataset 1.}  Features include: (\textbf{A. Correlation of split
  halves by number of patients included.})  For 500 iterations, we randomly split the
patient-specific correlations in half, then created one correlation
matrix with the first half and created the second correlation matrix
using more and more patients. The two correlation matrices were then correlated, and the
  coefficient plotted relative to the number of patients included.
  (\textbf{B. Standard error of the mean across iterations by the
    number of patients included.})
To measure the variability across the iterations, we plot the standard
error of the mean as a function of the number of patients included in
the second correlation matix.
  \textbf{C.--D. Features from Dataset 2.}  Analogous format to Panels
  A--B.}
\label{fig:supplemental_8}
\end{figure}

\clearpage
\newpage
\renewcommand{\bibnumfmt}[1]{[S#1]}
\renewcommand{\citenumfont}[1]{S#1}
\bibliography{CDL-bibliography/memlab}
\bibliographystyle{CSE}
% \newpage
% \renewcommand{\refname}{Supplemental references}
% \bibliography{memlab}

\end{document}